---
title: "Lab 4"
author: "Gaines"
date: "3/7/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
1. Consider the following example data for a between-subjects experiment with two groups, A and B:

example_data <- tibble(Group = rep(c("A","B"), each = 5),
                       DV = c(2,4,3,5,4,7,6,5,6,7))
                       
Use R to conduct a t.test and ANOVA on this data. Then use R to prove that the results of both analyses are the same. For example, prove that the p-values are the same, and prove that the F-value and T-value are related. (3 points)
```{r}
library(tidyverse)
example_data <- tibble(Group = rep(c("A","B"), each = 5),
                       DV = c(2,4,3,5,4,7,6,5,6,7))
t_object <- t.test(DV~Group, data = example_data, var.equal=TRUE)
aov_object <- aov(DV~Group, data=example_data)

class(t_object)
class(aov_object)
print(t_object)
sum.aov <- summary(aov_object)
##are f values and p values the same?
round(t_object$p.value, digits = 10) == round(sum.aov[[1]]$'Pr(>F)'[1], digits = 10)
##are the F and T values the same?
round(((t_object$statistic)^2), digits = 10) == round((sum.aov[[1]]$'F value'[1]), digits = 10)
```

```{r}
#explore how to use papaja to report analysis using example data
library(papaja)
apa_print_out <- papaja::apa_print(aov_object)
apa_print_out$estimate
apa_print_out$full_result
```
The main effect of the group was significant, `r apa_print_out$full_result`


2. Look at the lab on ANOVA that I wrote for our undergraduate statistics OER lab manual https://crumplab.github.io/statisticsLab/lab-8-one-way-anova.html. That lab shows an example of obtaining data from a published paper in psych science where a one-factor ANOVA was used as a part of the analysis. Load the data, conduct the ANOVA, report a ggplot of the means, and use papaja to help you write a short results section reporting the ANOVA result. (3 points).

```{r}
library(data.table)
library(papaja)
##Get data

library(readr)

all_data <- read_csv('https://raw.githubusercontent.com/CrumpLab/statisticsLab/master/data/Jamesetal2015Experiment2.csv')
all_data <- data.table(all_data)
all_data$Condition <- as.factor(all_data$Condition)
levels(all_data$Condition) <- c("Control",
                                "Reactivation+Tetris", 
                                "Tetris_only",
                                "Reactivation_only")
summary(aov(Days_One_to_Seven_Number_of_Intrusions ~ Condition, all_data))
aov_object_tetris <- aov(Days_One_to_Seven_Number_of_Intrusions ~ Condition, all_data)
aov_object_tetris_print <- papaja::apa_print(aov_object_tetris)
```


The effect of weird tetris therapy was significant, `r aov_object_tetris_print$full_result`

```{r}
ggplot(all_data, aes(x=Condition, 
                     y=Days_One_to_Seven_Number_of_Intrusions, 
                     fill = Condition))+
  geom_bar(stat="summary", fun = "mean")
```
